---
title: "regeneration_16s_part3"
output: html_document
date: "2024-07-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r packages composition}
library(rlang)
library(stringr)
library(dplyr)
library(stats)
library(ggpubr)
library(vegan)
library(cowplot)
library(tidyverse)
#library(MCMC.OTU)
#install.packages("remotes")
#remotes::install_github("Jtrachsel/funfuns")
library(phyloseq)
setwd("/projectnb/coral/jk_data/proj4_Regeneration/1_Raw_reads/Stella/")
```

Read in data

```{r read in data}
load("taxa2.Rdata")

#phyloseq objects
load("ps.clean.Rdata")
load("ps.rare.Rdata")
load("ps.trim.rare.Rdata") # rare otu's trimmed, then rarefied (1k)
ps.cleanest = readRDS("phyloseq.cleanest.all.rds")

# remove treatment NA's
ps.clean <- subset_samples(ps.clean,(!is.na(treatment)))
ps.rare <- subset_samples(ps.rare,(!is.na(treatment)))
ps.trim.rare <- subset_samples(ps.trim.rare,(!is.na(treatment)))


# moving forward with rarefied data, doesn't seem to make a difference in the results whether rare OTUs trimmed or not 

# make the factors of interest actual factors 
sample_data(ps.rare)$treatment = as.factor(sample_data(ps.rare)$treatment)
#levels(sample_data(ps.rare)$treatment)
# "Antibiotic" "Control"   


# set color palettes
colvec<- c("Control" = "#4292c6", "Antibiotic" = "coral1")
```

Rename ASVs to be more informative

```{r rename ASVs}
tax <- as.data.frame(ps.rare@tax_table@.Data)

tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "D_0__",""),
                        Phylum = str_replace(tax[,2], "D_1__",""),
                        Class = str_replace(tax[,3], "D_2__",""),
                        Order = str_replace(tax[,4], "D_3__",""),
                        Family = str_replace(tax[,5], "D_4__",""),
                        Genus = str_replace(tax[,6], "D_5__",""),
                        Species = str_replace(tax[,7], "D_6__",""),
                        stringsAsFactors = FALSE)
tax.clean[is.na(tax.clean)] <- ""

for (i in 1:7){ tax.clean[,i] <- as.character(tax.clean[,i])}
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
    kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Class_", tax.clean[i,3], sep = "")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Order_", tax.clean[i,4], sep = "")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Family_", tax.clean[i,5], sep = "")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Genus",tax.clean$Genus[i], sep = "_")
  }
}

tax_table(ps.rare) <- as.matrix(tax.clean)
```


# All ASVs - PCAs

## Rarefied Data

```{r }

ord <- ordinate(ps.rare, "PCoA", "bray")
library(ggplot2)
library(phyloseq)

pcoa.treat=plot_ordination(ps.rare, ord, color = "treatment") +
  geom_point(size = 0.12, alpha = 0.1) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = ps.rare@sam_data$treatment),lwd = 0.5, size = 0.5) +
  theme_classic()
pcoa.treat
ggsave(pcoa.treat, filename = "Stella_pcoas.allasvs.rare.pdf", width=3.5, height=2.5, units=c("in"), useDingbats=FALSE)

```

### Rarefied Stats

Help on adonis (here)[https://thebiobucket.blogspot.com/2011/04/assumptions-for-permanova-with-adonis.html#more]

```{r stats rarefied data}
seq.rare <- data.frame(otu_table(ps.rare))

# make distance matrix with vegdist():
dist.rare <- vegdist(seq.rare)
samdf.rare <- data.frame(sample_data(ps.rare))
row.names(samdf.rare)==row.names(seq.rare)

# Load required libraries
library(phyloseq)
library(vegan)  # For the vegdist() function

dist.rare_matrix <- as.matrix(dist.rare)

# Subset the distance matrix for the filtered samples
filtered_dist_matrix <- dist.rare_matrix[rownames(samdf.rare), rownames(samdf.rare)]

# Create a distance object from the filtered distance matrix
filtered_dist <- as.dist(filtered_dist_matrix)

# Perform beta dispersion analysis based on 'treatment'
bet.all <- betadisper(filtered_dist, samdf.rare$treatment)

# Perform ANOVA test
anova_results <- anova(bet.all)

# Print the ANOVA results
print(anova_results)
#p=0.3321
adonis2(seq.rare ~ treatment, data=samdf.rare, permutations=999)
#     Df SumOfSqs      R2      F Pr(>F)   
#treatment  1  0.22509 0.46869 6.1749  0.005 **
##Residual   7  0.25517 0.53131                 
#Total      8  0.48026 1.00000      
adonis2(formula = seq.rare ~ treatment, data = samdf.rare, permutations = 999)
#  treatment  1  0.22509 0.46869 6.1749  0.008 **
#Residual   7  0.25517 0.53131                 
#Total      8  0.48026 1.00000    

```
```{r}
library(phyloseq)
load(file = "ps.trim.Rdata")
phylo_trim_rel <- transform_sample_counts(ps.trim, function(x) x / sum(x))
phylo_trim_rel_df <- data.frame(tax_table(ps.trim))
library(phyloseq)
library(ggplot2)
library(RColorBrewer)
library(plotly)

plot_bar(phylo_trim_rel, x = "sample", fill="Phylum") +
  geom_bar(stat="identity") +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom") +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(11))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(11))) +
  ylab("Relative abundance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
## Selecting just Phylum Proteobacteria 
phylo_proteo <- subset_taxa(ps.trim, Phylum %in% "Proteobacteria")
phylo_proteo_rel <- transform_sample_counts(phylo_proteo, function(x) x / sum(x))

Proteo<-plot_bar(phylo_proteo_rel, x = "sample", fill = "Order")+
  geom_bar(stat="identity") +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(15))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(15))) +
  ylab("Relative abundance") +
  ggtitle("Proteobacteria ASVs (relative abundance)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
Proteo
ggsave(Proteo, filename = "Stella_Proteo.rare.pdf", width=3.2, height=4.5, units=c("in"), useDingbats=FALSE)
```

# Core vs. accessory

## Core Microbiome

```{r packages core, echo=FALSE}
#nBiocManager::install("microbiome")
#remotes::install_github("r-lib/rlang")
library(microbiome)
```

```{r core 15 taxa and 10 samples}
# prevalence here means what percent of samples does the taxa need to be in to be considered 'core' 

# Transform to compositional abundances
pseq.rel <- microbiome::transform(ps.rare, "compositional")
# Calculate prevalences for all taxonomic groups
head(prevalence(ps.rare, detection = 1/100, sort = TRUE))

# Pick the core (>0.1% relative abundance in >50% of the samples)
pseq.core <- core(ps.rare, detection = 0.1/100, prevalence = 0.5)
pseq.core
#19 taxa and 9 samples
seq.rare <- data.frame(otu_table(ps.rare))
#saving
core.tax <-data.frame(otu_table(pseq.core))
# Plot bar plots
ps_glom <- tax_glom(pseq.core, "Genus")
ps <- transform_sample_counts(ps_glom, function(x) x / sum(x))
ps1 <- merge_samples(ps, "sample")
ps2 <- transform_sample_counts(ps1, function(x) x / sum(x))
library(RColorBrewer)


plot=plot_bar(ps, x = "sample", fill = "Genus")+
  geom_bar(stat="identity") +
  theme_bw() +
  theme(panel.grid = element_blank(),legend.position = "bottom", legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(17))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(17))) +
  ylab("Relative abundance") +
  ggtitle("CORE ASVs (relative abundance)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot
ggsave(plot, file="Stella_core_rare.bar_1.pdf",width=3.5, height=2.9, units=c("in"), useDingbats=FALSE)

# Core Bray-Curtis PCoAs
#not rel abun
plot.core=plot_ordination(pseq.core,ordinate(pseq.core,"PCoA", "bray"),color="treatment")+
  stat_ellipse(level = 0.73)+
  theme_classic()+scale_color_manual(values = colvec) 
  
  #                   values=cols_site_diverging,
  #                   breaks = c("Control","Nitrate","Heat","Heat+Nitrate"))+
  #scale_shape_manual(name = "state", 
  #                   values=c(15,24),
  #                   breaks = c("apo","sym"))
plot.core
ggsave(plot.core, file="Stella_core_pca.pdf",width=4,height = 4)

seq.core <- data.frame(otu_table(pseq.core))
samdf.core <- data.frame(sample_data(pseq.core))
row.names(samdf.core)==row.names(seq.core)
#rel abundance
pseq.core.rel <- transform_sample_counts(pseq.core, function(x) x / sum(x))
# pcoa's 
plot.core.rel = plot_ordination(pseq.core.rel,ordinate(pseq.core.rel,"PCoA", "bray"),color="treatment")+
  stat_ellipse(aes(color = treatment))+
theme_classic()+scale_color_manual(values = colvec)
plot.core.rel
adonis2(core.tax ~ treatment, data=samdf.core, permutations=999)
#Df SumOfSqs      R2      F Pr(>F)   
#treatment  1  0.18994 0.50422 7.1191  0.007 **
#Residual   7  0.18676 0.49578                 
#Total      8  0.37670 1.00000         

#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
adonis2(formula = core.tax ~ treatment, data = samdf.core, permutations = 999)
ggsave(plot.core.rel, filename = "Stella_pcoas.core.rel.rare_pca.pdf", width=4, height = 3, units=c("in"), useDingbats=FALSE)
```

```{r}
# Assuming 'ps' is your phyloseq object
# Transform the counts to relative abundance
ps_rel <- transform_sample_counts(ps, function(x) x / sum(x))

# Extract data for plotting
ps_df <- psmelt(ps_rel)
x=Genus, y=log2FoldChange, fill=Phylum
# Convert 'Genus' to a factor to maintain order
ps_df$Phylum <- factor(ps_df$Phylum, levels = rev(names(sort(tapply(ps_df$Abundance, ps_df$Phylum, sum)))))
# Bubble plot
BUB_core <- ggplot(ps_df, aes(x = Sample, y = Class, size = Abundance, color =Phylum)) +
  geom_point(alpha = 1) +  # Add bubbles with some transparency
  theme_bw() +  # Use a clean white background
  theme(panel.grid = element_blank(),  # Remove grid lines
        legend.key.size = unit(0.4, 'cm'),  # Adjust legend key size
        axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for better readability
  scale_color_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(14))) +  # Set custom colors for the 'Genus'
  scale_size_continuous(range = c(1,9), name = "Relative Abundance") +  # Adjust bubble size range
  labs(y = "Class", x = "Sample", title = "CORE ASVs (Relative Abundance)")  # Set labels and title

# Print the plot
BUB_core
ggsave(BUB_core, file="Stella_core_rare.bar_1_BOB_ord.pdf",width=5.5,height = 3.4)

```


```{r calculate core abundance}
# calculating core abundances #
library(vctrs)
library(dplyr)
library(dplyr)
library(tidyselect)
core.sqs<- tax_table(pseq.core)
core.sqs.ids= row.names(core.sqs)
library(dplyr)
core.sqs.ids
# Assuming seq_rare_rel is your data frame or tibble
core_sqs_ids <- c("sq1","sq2","sq3","sq4","sq5","sq6","sq8","sq9","sq10","sq11","sq13","sq16","sq18","sq19","sq22","sq27","sq28","sq29","sq30")
seq.core <- seq_rare_rel[, core_sqs_ids]

core.rel <- data.frame(colMeans(seq.core))

total.rel <- data.frame(colMeans(seq.rare.rel))
total.rel.ordered <- data.frame(total.rel[order(-total.rel$colMeans.seq.rare.rel.),,drop=FALSE])

```

### Core stats

```{r core stats Lineages together}
seq.core <- data.frame(otu_table(pseq.core))

dist.core <- vegdist(seq.core)
samdf.core <- data.frame(sample_data(pseq.core))
row.names(samdf.core)==row.names(seq.core)

# site
bet.all <- betadisper(dist.core,samdf.core$treatment)
anova(bet.all) # 0.3529
permutest(bet.all, pairwise = TRUE, permutations = 999)
      
#Df   Sum Sq   Mean Sq      F N.Perm Pr(>F)
#Groups     1 0.021791 0.0217906 2.8006    999   0.12
#Residuals  8 0.062247 0.0077808 

```

## Accessory

```{r accessory lineages together}
# not doing prestress data here because we only have 4 taxa for prestress data
ps.rare.otu <- data.frame(ps.rare@otu_table)
core.tax <- data.frame(pseq.core@tax_table)
core.ids <- c(rownames(core.tax))
ps.rare.acc.otu <- ps.rare.otu[,!colnames(ps.rare.otu) %in% core.ids ]
row.names(samdf.core) <- samdf.core$sample

#remake phyloseq object
ps.acc <- phyloseq(otu_table(ps.rare.acc.otu, taxa_are_rows=FALSE), 
                         sample_data(samdf.core), 
                         tax_table(taxa2))
ps.acc #82 taxa and 9 samples


# make bray-curtis dissimilarity pcoas for accessory microbiome

pcoa.acc.all=plot_ordination(ps.acc,ordinate(ps.acc,"PCoA", "bray"), color = "treatment") +
  geom_point(size = 2.5, alpha = 0.8) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = ps.rare@sam_data$treatment), type = "t", lwd = 1,level = 0.79) +
  theme_classic()
pcoa.acc.all

ggsave(pcoa.acc.all, filename = "Stella_pcoas.accesory.rel.rare.pdf", width=4, height = 3, units=c("in"), useDingbats=FALSE)
#rel abundance
pseq.acc.rel <- transform_sample_counts(ps.acc, function(x) x / sum(x))

```



```{r stats acc 85 taxa and 10 samples}
seq.acc <- data.frame(otu_table(ps.acc))

dist.acc <- vegdist(seq.acc)
samdf.acc <- data.frame(sample_data(ps.acc))
row.names(samdf.acc)==row.names(seq.acc)

#sitename
bet.all <- betadisper(dist.acc,samdf.acc$treatment)
anova(bet.all) #p=0.3568
#Df   Sum Sq  Mean Sq F value   Pr(>F)   
#Groups     1 0.035551 0.035551  12.731 0.009121 **
#Residuals  7 0.019547 0.002792 
permutest(bet.all, pairwise = TRUE, permutations = 999) 
plot(bet.all) 
#Df   Sum Sq  Mean Sq      F N.Perm Pr(>F)    
#Groups     1 0.035551 0.035551 12.731    999  0.001 ***
#Residuals  7 0.019547 0.002792
adonis2(seq.acc ~ treatment, data=samdf.acc, permutations=999) 
# Df SumOfSqs      R2      F Pr(>F)   
#treatment  1  0.62771 0.20095 1.7605  0.007 **
#Residual   7  2.49593 0.79905                 
#Total      8  3.12364 1.00000               
```



## Accessory Summary

# Bar plots {.tabset}
```{r}
# Pick the core (>0.1% relative abundance in >50% of the samples)

seq.acc <- data.frame(otu_table(ps.acc))
# Plot bar plots


# Transform to compositional abundances
pseq.acc <- microbiome::transform(ps.acc, "compositional")
# Calculate prevalences for all taxonomic groups
head(prevalence(ps.rare, detection = 1/100, sort = TRUE))

#15 taxa and 10 samples
seq.acc <- data.frame(otu_table(pseq.acc))
# Plot bar plots
ps_glom <- tax_glom(ps.acc, "Genus")
ps <- transform_sample_counts(ps_glom, function(x) x / sum(x))
ps1 <- merge_samples(ps, "sample")
ps2 <- transform_sample_counts(ps1, function(x) x / sum(x))
library(RColorBrewer)


ps_glom_1 <- tax_glom(seq.acc, "Genus")
ps <- transform_sample_counts(ps_glom, function(x) x / sum(x))
ps1 <- merge_samples(ps, "sample")
ps2 <- transform_sample_counts(ps1, function(x) x / sum(x))
library(RColorBrewer)

plot=plot_bar(ps, x = "sample", fill = "Class")+
  geom_bar(stat="identity") +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(14))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(14))) +
  ylab("Relative abundance") +
  ggtitle("Accessory ASVs (relative abundance)")+theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot
ggsave(plot, file="Stella_acc_rare.bar_1.pdf",width=7.5, height=2.9, units=c("in"), useDingbats=FALSE)
```
```{r}
ps_no_gamma <- subset_taxa(ps, Class != "Gammaproteobacteria")
ps_no_gamma
ps1 <- merge_samples(ps_no_gamma, "sample")
ps2 <- transform_sample_counts(ps_no_gamma, function(x) x / sum(x))
library(RColorBrewer)
ps2
plot <- plot_bar(ps2, x = "sample", fill = "Class") +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(14))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(14))) +
  ylab("Relative abundance") +
  ggtitle("Accessory ASVs (relative abundance)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot)

```


```{r}
# Assuming ps is your phyloseq object
# Transform the counts to relative abundance
ps_rel <- transform_sample_counts(ps, function(x) x / sum(x))

# Extract data for plotting
ps_df <- psmelt(ps_rel)

# Convert `Genus` to a factor with the desired order
ps_df$Genus <- factor(ps_df$Genus, levels = rev(names(sort(tapply(ps_df$Abundance, ps_df$Genus, sum)))))

```

```{r}
# Bubble plot
BUB<-ggplot(ps_df, aes(x = sample, y = Genus, size = Abundance, color = Genus)) +
  geom_point(alpha = 0.7) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.key.size = unit(0.3, 'cm')) +
  scale_color_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(length(unique(ps_df$Genus))))) +
  scale_size_continuous(range = c(1, 10), name = "Relative Abundance") +
  labs(y = "Genus", x = "Sample", title = "Accessory ASVs (Relative Abundance)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Optional: Rotate x-axis labels for better readability
BUB
ggsave(BUB, file="Stella_acc_rare.bar_1_BOB.pdf",width=14)
```



## Summary

- No differences in results between relative abundance & rarefied plots/stats
- Pairwise adonis significant between control treatment and all other variability treatments
- No lineage effects to suggest cryptic lineages host different microbiomes
- Beta dispersion not significantly different 

# Pgylosq Deseq
#Detect differentially abundant taxa

Tutorial [here](https://joey711.github.io/phyloseq-extensions/DESeq2.html)

## Setup
```{r}
#Use unrarefied data for this analysis. I used the clean table (contaminant removed)
library("DESeq2")
packageVersion("DESeq2")
setwd("/projectnb/coral/jk_data/proj4_Regeneration/1_Raw_reads/Stella")
```


```{r}
diagdds = phyloseq_to_deseq2(ps.cleanest, ~ treatment)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
```

```{r}
#calculate foldchange. We put the experimental first and control second. 
res = results(diagdds, cooksCutoff = FALSE)
res = res[order(res$pvalue, na.last=NA), ]
sigtab05 = res[(res$padj < 0.1), ]
dim(sigtab05)
sigtab05
write.csv(sigtab05, "/projectnb/coral/jk_data/proj4_Regeneration/1_Raw_reads/Stella/Stella_diff_sig.csv", row.names = TRUE)
```

```{r}
#calculate foldchange. We put the experimental first and control second. 
# Extract results for Antibiotic vs Control
res <- results(diagdds, contrast = c("treatment", "Antibiotic", "Control"))

# Inspect the results
head(res)

# Save results to a dataframe for further analysis or plotting
res<- as.data.frame(res)
res = res[order(res$pvalue, na.last=NA), ]
sigtab05 = res[(res$pvalue < 0.05), ]
dim(sigtab05)
sigtab05
write.csv(sigtab05, "/projectnb/coral/jk_data/proj4_Regeneration/1_Raw_reads/Stella/Stella_diff_sig.csv", row.names = TRUE)
```

```{r}
#Export this to use as a table (for supplementary)
sigtab05 = cbind(as(sigtab05, "data.frame"), as(tax_table(ps.clean)[rownames(sigtab05), ], "matrix"))
#First, asting the factors to characters so they can be replaced..
sigtab05$Genus <-as.character(sigtab05$Genus)
sigtab05
```
```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ANCOMBC")
library(ANCOMBC)
```

```{r}
#Define our discrete palette
#scale_fill_discrete <- function(palname = "Set1", ...) {
#scale_fill_brewer(palette = palname, ...)
#}
#Sort our table
x = tapply(sigtab05$log2FoldChange, sigtab05$Phylum, function(x) max(x))
x = sort(x, TRUE)
#Cast as factors
sigtab05$Phylum = factor(as.character(sigtab05$Phylum), levels=names(x))
#Same for genus
x = tapply(sigtab05$log2FoldChange, sigtab05$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab05$Genus = factor(as.character(sigtab05$Genus), levels=names(x))
#Count how many NAs increase and decrease in abundance, make a note of it and them omit them for plotting. 
sigtab05pos = sigtab05[(sigtab05$log2FoldChange >=0),]
pos.row.has.na <- apply(sigtab05pos, 1, function(x){any(is.na(x))})
sum(pos.row.has.na)
```
```{r}
sigtab05neg = sigtab05[(sigtab05$log2FoldChange <=0),]
neg.row.has.na <- apply(sigtab05neg, 1, function(x){any(is.na(x))})
sum(neg.row.has.na)
```

```{r}
#Filter these NAs fro mthe table we will use for plotting..
row.has.na <- apply(sigtab05, 1, function(x){any(is.na(x))})
sigtab05.filtered <- sigtab05[!row.has.na,]
sigtab05.filtered
```
```{r}
library(ggplot2)
library(RColorBrewer)

ggplot(sigtab05, aes(x=Kingdom, y=log2FoldChange, fill=Kingdom)) + 
  geom_jitter(size=3, shape=21, width=0.3) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5), 
        panel.grid.minor = element_blank()) + 
  ylab("log2FoldChange") + 
  theme_classic() + 
  scale_fill_manual(values=brewer.pal(n = length(unique(sigtab05$Kingdom)), name = "Set1")) + 
  coord_flip()

```

```{r}
library(ggplot2)
library(RColorBrewer)
# Assuming `sigtab05.filtered` is your data frame
ggplot(sigtab05, aes(x=Kingdom, y=log2FoldChange, fill=Kingdom)) + 
  geom_jitter(size=3, shape=21, width = 0.3) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5), 
        panel.grid.minor = element_blank()) + 
  ylab("log2FoldChange") + 
  theme_classic()+
  scale_fill_manual(values=brewer.pal(n = length(unique(sigtab05$Kingdom), name = "Set1")) + 
  scale_color_manual(values=brewer.pal(n = length(unique(sigtab05$Kingdom)), name = "Set1"))+coord_flip()

# Additional customization for fill colors using scale_fill_hue
de# + scale_fill_hue(l=80, c=150)
ggsave(de, filename = "Stella_DF.pdf", width=5, height=1.5, units=c("in"), useDingbats=FALSE)
```
```{r}
# Load necessary libraries
library(phyloseq)
library(ALDEx2)

# Load your phyloseq object
phyloseq_object <- readRDS("phyloseq.cleanest.all.rds")
# Agglomerate taxa to order level
phyloseq_order <- tax_glom(phyloseq_object, taxrank = "Order")
# Transform the OTU table to relative abundances
phyloseq_order_rel <- transform_sample_counts(phyloseq_order, function(x) x / sum(x))
# Load the necessary library for plotting
library(ggplot2)

# Create a barplot of the orders
plot_bar(phyloseq_order_rel, fill = "Order") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Relative Abundance of Orders", y = "Relative Abundance", x = "Sample")

# Extract the OTU table and metadata from the phyloseq object
otu_table_order <- as(otu_table(phyloseq_order), "matrix")

# If the OTU table has sample names as row names, transpose it
otu_table_order <- t(otu_table_order)
tax_table_order <- as(tax_table(phyloseq_order), "matrix")
# Extract the sample metadata from the phyloseq object and convert it to a data frame
metadata <- as(sample_data(phyloseq_order), "data.frame")

# Ensure that the metadata contains a column for the group (e.g., "treatment" with "antibiotic" and "control")
group_variable <- "treatment"  # Change this to your actual group variable name
# Transform the OTU table using center-log ratio (CLR) transformation
clr_data_order <- aldex.clr(otu_table_order, metadata[[group_variable]], mc.samples=128, denom="all")

# Run ANCOM
ancom_results_order <- aldex.kw(clr_data_order)
View(ancom_results_order)

# Filter significant results
significant_results_order <- ancom_results_order[ancom_results_order$we.eBH < 0.9, ]
print(significant_results_order)
# Extract counts for Burkholderiales
burkholderiales_counts <- otu_table_order[, tax_table_order[, "Order"] == "Burkholderiales"]
```

```{r}
# Transpose OTU table if needed
otu_table_order <- as(otu_table(phyloseq_order), "matrix")
otu_table_order <- t(otu_table_order)

# Extract the taxonomy table
tax_table_order <- as(tax_table(phyloseq_order), "matrix")
# Identify the taxa that belong to the order "Burkholderiales"
burkholderiales_taxa <- rownames(tax_table_order)[tax_table_order[, "Order"] == "Burkholderiales"]
sq3_counts <- otu_table_order["sq3", ]
print(sq3_counts)
# Assuming you have a metadata dataframe with sample names as rownames
metadata <- as(sample_data(phyloseq_order), "data.frame")

# Convert sq3_counts to a data frame and add sample names as a column
sq3_counts_df <- as.data.frame(sq3_counts)
sq3_counts_df$Sample <- rownames(sq3_counts_df)

# Merge with metadata
merged_data <- merge(sq3_counts_df, metadata, by.x = "Sample", by.y = "row.names")

# Example: T-test between two groups
t_test_result <- t.test(sq3_counts ~ merged_data$treatment)

# Print the t-test results
print(t_test_result)
# Create the boxplot
ggplot(merged_data, aes(x = treatment, y = sq3_counts)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5) +  # Optional: Add jitter to see individual points
  labs(title = "Counts of sq3 in Control vs. Antibiotic Groups", 
       x = "Treatment", 
       y = "Counts of sq3") +
  theme_minimal()
```
```{r}
# View unique entries in the "Order" column
# Extract taxa that belong to Burkholderiales
burkholderiales_taxa <- rownames(tax_table_order)[tax_table_order[, "Order"] == "Burkholderiales"]


```



